---
title: "01_cal07_renamed.analysis"
output: html_document
---
Analysis based primarily on the Sun and Vera et. al. 2020 paper with the code from https://github.com/BROOKELAB/SingleCell/tree/fa170e4c16f64670bada6ef56ae9cf1ffafb4786 and the Seurat database vignettes https://github.com/satijalab/seurat/blob/master/vignettes/pbmc3k_tutorial.Rmd  
Preliminary analysis with regressing out the cell cycle markers that skew the clustering

*This document can't be knit yet as it throws an error at line 156 and I need to figure out how to fix that

```{r}
# load required libraries
library(Seurat)
library(ggplot2)
library(sctransform)
library(dplyr)
```

Read in the .h5 file using Seurat function. Load in the count matrix as a Seurat object. 
```{r}
# read in Cal07 filtered data
cal07_a549_counts <- Read10X_h5(
        "/Users/ethayer/Google Drive/Grad School/Brooke Lab/Data/polyIC_scRNAseq/03_scRNAseq_h5/Cal07_aggr/filtered_feature_bc_matrix.h5",
        use.names = TRUE, unique.features = TRUE)
# load the data into Seurat for analysis
a549 <- CreateSeuratObject(counts = cal07_a549_counts, min.cells = 4, 
                           min.features = 400, project = "cal07_mock")
```
Organize the data into two levels based off of the barcode ids (in this case Cal07(1) or Mock(2)). 
```{r}
# split the rownames by -, then select only the second element 
# (the number 1 (cal07) or 2 (mock))
experiment_id <- factor(sapply(strsplit(rownames(a549@meta.data), "-"), 
                               function(x) x[2]))
# set the factor levels
levels(experiment_id) <- c("Cal07", "Mock")

# Add it back to the seurat object meta-data
a549@meta.data$experiment_id <- experiment_id
```


Look at QC metrics of the data. 
```{r}
# Show QC metrics for the first/last 20 cells
head(a549@meta.data, 20)
tail(a549@meta.data, 20)
summary(a549@meta.data$experiment_id)

# Visualize QC metrics as a violin plot
VlnPlot(a549, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
        
# Look at the relationship between counts and features
FeatureScatter(a549, feature1 = "nCount_RNA", feature2 = "nFeature_RNA",
               group.by =  "experiment_id")
```

Filtering based on QC. Low feature could mean this is a droplet. High count/feature indicates a doublet.
```{r}
# Subset to a new object based off of the QC data 
a549_01 <- subset(a549, 
                  subset = nFeature_RNA > 500 & nFeature_RNA < 8500 & nCount_RNA < 115000)

expr <- FetchData(object = object, vars = var1)
object[, which(x = expr > low & expr < high)]
a549_01@meta.data
# Do all the QC stuff again with the new object
head(a549_01@meta.data, 20)
# Visualize QC metrics as a violin plot
VlnPlot(a549_01, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)
# Look at the relationship between counts and features
FeatureScatter(a549_01, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", 
               group.by =  "experiment_id")

```

Loading in the cell cycle markers from Tirosh et. al., 2015 with Seurat. We can then assign cell cycle scores.
```{r}
# Loading in the markers
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

# Assign cell cycle scores based off of the imported cycle markers above
a549_01 <- CellCycleScoring(
        a549_01,
        s.features = s.genes,
        g2m.features = g2m.genes,
        set.ident = TRUE)

# view cell cycle scores and phase assignments
head(a549_01[[]])
```


```{r}
# different scaling method which allows you to specify variables contributing to the scaling/normalization of the data
a549_01 <- SCTransform(
        a549_01, method = "glmGamPoi",
        vars.to.regress = c("S.Score", "G2M.Score"),
        verbose = FALSE)
```

Looking at which dimensions contribute the most to the percentage of variance
```{r}
# PCA to find similar genes/cells for clustering
a549_01 <- RunPCA(a549_01, verbose = FALSE)
# Best explanation from https://github.com/satijalab/seurat/blob/master/vignettes/pbmc3k_tutorial.Rmd:
# "‘Elbow plot’: a ranking of principle components based on the percentage of variance explained by each one"
# I chose 15, as there is little drop off after that and it is best to be conservative when picking your dimensions to not leave anything out
ElbowPlot(a549_01)
```

Performing dimensional reduction and clustering
```{r}
# making UMAP dimensional reduction
a549_01 <- RunUMAP(a549_01, dims = 1:15, verbose = FALSE)
a549_01 <- FindNeighbors(a549_01, dims = 1:15, verbose = FALSE)
a549_01 <- FindClusters(a549_01, verbose = FALSE, resolution = 0.5)
DimPlot(a549_01, label = TRUE) + NoLegend()
```

Want to make sure the cell cycle genes are regressed out
```{r}
# loop to plot each UMAP with colored cell cycle genes one by one
for (x in g2m.genes[g2m.genes %in% VariableFeatures(a549)]) {
    g <- FeaturePlot(a549_01, x, order = TRUE, pt.size = 1)  
    print(g)
}

# UMAP color/clustering based on classification of cell cycle
# making sure the cell cycle genes are widely distributed
FeaturePlot(a549_01, "G2M.Score")
```

Identifying the most variable genes
```{r}
# identify features that are outliers on a 'mean variability plot'
a549_01 <- FindVariableFeatures(a549_01, selection.method = "vst", nfeatures = 2000)

# Identify the 25 most highly variable genes
top25 <- head(VariableFeatures(a549_01), 15)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(a549_01)
plot2 <- LabelPoints(plot = plot1, points = top25, repel = TRUE,
                     max.overlaps = 20, xnudge = 0, ynudge = 0)
plot2
```

Where in the clustering are the viral genes being expressed
```{r}
# where are the infected cells in the dimensional reduction
segment_gene_names <- c("NS", "HA", "NA", "NP", "PB2", "PB1", "PA", "M")

# loop to plot each UMAP with colored flu genes one by one
for (x in segment_gene_names) {
    g <- FeaturePlot(a549_01, x, order = TRUE, pt.size = 1)  
    print(g)
}
```

# Calculating the viral score

Still needed to do: 
- set cutoffs for each viral gene based off the first local minimum
- see what % of the population is infected
- look at IFN & ISG landscape

Organizing to just look at  Cal07-exposed population (hopefully infected)
```{r}
# subetting to select only cells from the Cal07 population
just_cal <- subset(x = a549_01, subset = experiment_id == "Cal07")
just_cal
```

Setting cutoffs and quickly trying to look at genes... ending here for now
```{r}
# This will calculate the number of UMI that match flu segments in each cell
# starting with the subset that only has the cal07 population
flu_counts <- colSums(GetAssayData(just_cal, slot = "count")[segment_gene_names,])
all_counts <- colSums(GetAssayData(just_cal, slot = "count"))
# calculating the percentage of flu in population
just_cal$flu_pct_count <- flu_counts / all_counts
just_cal$log_flu_pct_count <- log10((just_cal$flu_pct_count * 100) + 1)

# This will calculate the number of UMI that match flu segments in each cell 
# this is the aggregated object
flu_counts2 <- colSums(GetAssayData(a549_01, slot = "count")[segment_gene_names,])
all_counts2 <- colSums(GetAssayData(a549_01, slot = "count"))

a549_01$flu_pct_count2 <- flu_counts2 / all_counts2
a549_01$log_flu_pct_count2 <- log10((a549_01$flu_pct_count2 * 100) + 1)

# getting a density line from the subset data
des.all <- density(just_cal$log_flu_pct_count)
# finding first local minima and setting as the cutoff
min.all <- des.all$x[which(diff(sign(diff(des.all$y)))==2)+1]
min.all

# density plot of aggregated data with the cutoff from subset
qplot(x = a549_01$log_flu_pct_count2, geom = "density") + scale_y_sqrt() +
        geom_vline(xintercept = min.all[1])

FeaturePlot(a549_01, "log_flu_pct_count")

a549_01$is_infected <- a549_01$log_flu_pct_count2 > min.all[1]
FeaturePlot(a549_01, "is_infected")
FeaturePlot(a549_01, "is_infected", split.by = "experiment_id", keep.scale = "all")
FeaturePlot(a549_01, "SLFN5", split.by = "experiment_id", keep.scale = "all")

VlnPlot(a549_01, "SLFN5", split.by = "is_infected", group.by = "experiment_id")
grep("SLFN5", VariableFeatures(a549_01), value = TRUE)

Idents(a549_01) <- a549_01$is_infected
FindAllMarkers(a549_01)
```

